---
title: "Day 4 - Comp session"
format: html
editor: visual
---
```{r}
library(tidyverse)
```

```{r}
quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_time)
}
```

## Intro to functions

Make a function to double the user input.

```{r}
double_this <- function (my_value) {
  print(2 * my_value)
}

double_this(my_value = 200)
double_this(200)

double_this(c(1, 2, 10, 100, 175))
```

Your turn: Write a function that asks for a user's name and age and returns an exclamation, like "I'm \[name\], and I am \[age\] years old."

```{r}
exclaim_name_and_age <- function(name, age) {
  print(paste("I'm", paste0(name, ","), "and I am", age, "years old!"))
}

exclaim_name_and_age("Erika", 22)
```

```{r}
convert_animal_age <- function(animal, age) {
  if (animal == "dog") {
    print(age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  } else {
    stop("Not a valid animal choice")
  }
}

convert_animal_age(animal = "dog", age = 8)
convert_animal_age(animal = "goat", age = 8)
convert_animal_age(animal = "parrot", age = 8)
```

```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$food))
}

dog_menu("Waffle")
```

## Functions meet for loops

```{r}
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)

means_over_columns(df = faithful)
```

## Practice
```{r}
animals <- c("orcas","band-rumped storm petre", "monarch butterfly","snow leopards" )
territories <- c("salish sea","hawai'i", "mazahua","Buryat")
 
animal_territory <- matrix(nrow=length(animals), ncol=length(territories))

for (i in seq_along(animals)) {
  for (j in seq_along(territories)) {
    joined <- paste(territories[j],animals[i])
    animal_territory[i, j] <- joined
  }
}

animal_territory
```

## Logistic growth

Modeling a populartion over time and varying growth rate

```{r}
my_function <- function(oranges, rainbows) {
  (oranges + rainbows / 2.4)**2
} # shortcut cmnd+opt+x
```

```{r}
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

# test
logistic_growth(5000, 100, 0.07, 42)
```

```{r}
time_vec <- seq(from = 0, to = 35, by = 0.1)

pop_35 <- logistic_growth(N0 = 40, K = 8000, r = 0.1, time = time_vec) # try r = .8

pop_time35_df <- data.frame(time_vec, pop_35)
```

```{r}
ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35)) +
  geom_line(size = 0.5)
```

```{r}
# to use a vector instead
pop_35_vec <- vector(mode = "numeric", length = length(time_vec))

for(i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time = time_vec[i])
  pop_35_vec[i] <- population
}
```

Now extend this to iterate over multiple growth rates

```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
```

```{r}
# Create a for loop that goes through each, apply the 
# logistic growth function for a range of times for each growth rate

# Need to create a MATRIX to store the outputs in: 
out_matrix <- matrix(nrow = length(time_vec), ncol = length(r_seq))

# Now, a nested for loop:
for (i in seq_along(r_seq)) { # Outer loop is the growth rates
  for (j in seq_along(time_vec)) { # Inner loop is the time sequence values
  pop <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
  out_matrix[j, i] <- pop # Store the value in the appropriate row & column
  } 
}
```

```{r}
# Let's wrangling it a little bit 
out_df <- data.frame(out_matrix, time = time_vec) # Make it a data frame and add time

# Update the column names of out_df, keeping time column name the same
colnames(out_df) <- c(paste0("gr_", r_seq), "time")

# pivot_longer to make it tidy (you'll learn more about this next week)
out_df_long <- out_df %>% 
  pivot_longer(cols = -time, names_to = "growth_rate", values_to = "population")

# Then plot it: 
ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate)) +
  theme_minimal()
```

## A function with warnings and an error message

Warning for a warning and stop to add an error

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if (windspeed > 130) {
    warning("Whoa, that's a high windspeed! Are you sure that's correct?")
  }
  
  if (rho > 1.225) {
    warning("That air density is suspicious.")
  }
  
  if (radius < 0) {
    stop("Rotor radius must be a positive value (meters).")
  }
  
  0.3*rho*pi*(radius^2)*(windspeed^3)
  
}

calc_windpower(2, -1, 131)
```

